---
title: "finViz Demo"
author: "James Del Mutolo"
date: "2025-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## finViz Package

The finViz package depends on the NetworkD3 package which contains functions for creating a Sankey diagram. A Sankey diagram can be used for visualizing the flow of money through different categories and merchants.

The functions in the finViz package are used on .csv files that must be organized in a certain manner. The files must contain columns for: Date, Merchant, Category, and Amount.

## Loading the finViz Package
The finViz package can be installed directly from GitHub using the install_github() function from devtools.

```{r, eval = FALSE}
devtools::install_github("delmutoloj/finViz")
```

## Datasets

Functions in the finViz package are passed names of csv files that are then read into r within the function. csv files can not be included in an R package in the same way .rda files are. Because of this, the .csv files must be downloaded separately.

```{r}
setwd("G:/My Drive/USF/R Programming/Week 12 - R Markdown")

# Read deposits example file
deposits <- read.csv("deposits.csv")


# List of statement data to pass to mergeStatementDF()
statements <- list("credit_statement1.csv",
                   "credit_statement2.csv",
                   "checking_statement.csv")
```

## Functions

**statementCombine()**

Reads and combines multiple financial statement .csv files into a single data frame. It ensures the Date column is properly formatted and orders the data chronologically.

```{r}
library(finViz)

# Read the csv files into R, and combine the statements.
spending <- statementCombine(statements)
```

**financialSummary()**

Generates a summary of financial data, including total income, total expenses, net balance, and aggregated amounts by category and merchant.

```{r}
# Print financial summary
financialSummary(deposits, spending)
```

**financialSankey()**

Generates an interactive Sankey diagram to visualize income allocation and spending patterns. It dynamically determines income categories from the deposit data and structures financial transactions into a hierarchical flow, showing how income is distributed across expense categories and vendors.

```{r, dpi = 150, out.width = "900px"}
# Create sankey diagram
sankey <- financialSankey(deposits, spending)
sankey
```